---
title: "Seward Park Hemlock & Woodpecker Study"
output: html_document
css: style.css
---


```{r setup, include = TRUE, echo=TRUE}
f <- "https://pshannon.net/hemlockSurvey2024/tblFinal.csv"
tbl <- read.table(f, sep=",", header=TRUE, as.is=TRUE, nrow=-1, quote="")
dim(tbl)
```

### The simplest question:  does woodpecker damage correlate with dbh? health status?


```{r simpleCors, include=TRUE, echo=TRUE}

cor(tbl$wpDamage, tbl$dbh)
cor(tbl$wpDamage, tbl$h, use="pairwise.complete.obs")
```

### Combined effects?

We now know that woodpecker damage is somewhat related to both dbh and health.
How well do these factors in combination predict wpDamage?

```{r lm, include=TRUE, echo=TRUE}
model <- lm(wpDamage ~ 0 + h + dbh, data=tbl)
summary(model) 
```

### Interpreting lm (linear model) results
Extract the coefficients from the model   
```{r interpret-lm, include=TRUE, echo=TRUE}
coef(model)
hist(resid(model))  # 
```

### include more factors.  are h and dbh still the best predictors?
```{r lm2, include=TRUE, echo=TRUE}
model <- lm(wpDamage ~ h + canopy + dfAssoc + aspect + slope + h + dbh, data=tbl)
summary(model) 
```
### Perhaps there is a density effect?  Constrain to hemlock graveyard boundaries.

```{r densities, include=TRUE, echo=TRUE}

library(MASS)
library(secr)
tbl.sick <- subset(tbl, h<=1)
cellCount <- 500
limits <- c(-122.25676, -122.248006, 47.548233, 47.562744) # open street map

f.graveyard <- kde2d(tbl.sick$lon, tbl.sick$lat, n=cellCount, lims=limits)

 # look at the successively more inclusive graveyard boundaries
contour(f.graveyard)

 # now get the data, construct a polygon we can use to
 # identify trees within the boundary
 
boundaries.graveyard <- contourLines(f.graveyard)
length(boundaries.graveyard) # 15
# find out how many points define each boundary
# this indicates the size of the boundary
for(i  in seq_len(length(boundaries.graveyard)))
   print(sprintf("%d) %d points in polygon", i, length(boundaries.graveyard[[i]]$x)))
   

# I know from earlier work that the level 10 boundary is useful, not too big
# not too small, and close to what we see by observation

contour.number <- 10
graveyard.poly <- data.frame(x=boundaries.graveyard[[contour.number]]$x,
                               y=boundaries.graveyard[[contour.number]]$y)

# how many points are in this region? 214 of the 724 
tbl.rows.in.boundary <- which(pointsInPolygon(tbl[, c("lon", "lat")], graveyard.poly))
print(length(which(pointsInPolygon(tbl[, c("lon", "lat")], graveyard.poly))))
# which tbl rows describe trees in this graveyard region?
tbl.sub <- tbl[tbl.rows.in.boundary,]
```
### Calculate the same correlations seen above: do dbh and health in the graveyard
# correlate better than across the entire dataset?

```{r}
cor(tbl.sub$wpDamage, tbl.sub$dbh)
cor(tbl.sub$wpDamage, tbl.sub$h, use="pairwise.complete.obs")
```
### Finally, look for combined linear effects of dbh and health in this subtable
Compare the pvalues "Pr(>|t|)" against those for the whole table, above
```{r lm3, include=TRUE, echo=TRUE}
model <- lm(wpDamage ~ 0 + h + dbh, data=tbl.sub)
summary(model) 
```




